---
title: "Analyze Mental health disorder in Tech Companies using R"
author: "Chitramoy Mukherjee"
date: "2023-02-16"
output: pdf_document
---

## Introduction :

In the document we are going to analyze, How does employees mental health 
affects employees performance and overall company productivity in tech 
companies. 

Mental health affects your emotional, psychological and 
social well-being.Mental health is a key factor to determine the productivity of
the employee as well as the total performance of the company. If someone is
not mentally fit, he can't produce the expected output what he is capable of and
it also impacts his co-workers performance and impacts the work environment.

This sort of analysis helps the employer to identify and support an individual 
who may be experiencing a mental health or substance use concern or crisis 
and connect them with the appropriate employee resources.This sort of analysis 
helps to increase mental health awareness in employee and co-workers. As a whole 
this sort of analysis helps employer to increase productivity and make sure they
are providing a healthy work environment and takes care of it's employees mental
health. This allows employer to recognize the signs of someone who maybe
struggling and teaches them the skills to know when to reach out and what
resources are available.Organizations that incorporate mental health awareness 
help to create a healthy and productive work environment that reduces the 
stigma associated with mental illness, increases the organizations mental 
health literacy and teaches the skills to safely and responsibly respond 
to a co-workers mental health concern.

This topic is relevant to data science as we can analyze and identify the 
factors/variables that impacts the mental health and justify the relations 
between variables which is closely related to determine the mental health of
employees.We can create a model and feed data into it to identify the employees
mental health in the company and provide directions to them to overcome the 
situation.

## Research questions :

1. no_employees and tech_company are statistically significant predictors 
   of mental health condition?
1. Does family_history have any relation with mentalhealth?
1. How does the frequency of mental health illness differ by workpace 
   type (i.e., tech vs. non-tech) and size? 
1. How easy is it for you to take medical leave for a mental health condition?
1. Is there any relationship between the responders location and the 
   occurrence of mental illness?
1. Would you bring up a mental health issue with a potential employer in an interview?
1. Is the number of employees in the work place a factor of causing mental 
   illness?
1. Which variable has a correlated relationship with the occurrence of mental 
   illness?
1. Would you be willing to discuss a mental health issue with your coworkers?
1. Does the care_options in tech companies varies based on country?

## Approach :

* First step for any analysis in data science is to view the data in the dataset
  and understand the structure of the imported data.
* Understand each attributes in the dataset.
* Data cleansing will be the next step and Null/NA data handling in the source
  dataset to make it ready for analysis.
* Outlier Analysis and Treatment.Aggregation functions are very useful for 
  understanding the data and present its summarized picture.
* Identify the dependent and independent variable from the dataset.
* perform plotting using the differnt library functions.
* Compare and contrast predictive models using simple linear, 
  multiple linear, and polynomial regression methods.
* Evaluate a model for overfitting and underfitting conditions and tune 
  its performance using regularization and grid search.
* Create training and test dataset from source data to perform accuracy 
  testing of the model.
  
## Approach to address the problem :

* Prepare data for analysis by handling missing values, formatting and 
  normalizing data, binning, and turning categorical values into numeric values.
* Examine data using descriptive statistics, data grouping, 
  analysis of variance (ANOVA), and correlation statistics.
* Compare and contrast predictive models using simple linear, multiple linear, 
  and polynomial regression methods.
* Evaluate a model for overfitting and underfitting conditions and 
  tune its performance using regularization and grid search.
  
## Datasets and Source of Data :

1. 3 datasets will be analyzed as a pert of this analysis which sourced from one
    dataset and splitted based on country.
    * survey_USA
    * survey_UK
    * survey_Others
    
1. Data sourced from : 
  [KAGGLE](https://www.kaggle.com/code/aditimulye/mental-health-at-workplace/data).
  
This dataset is from a 2014 survey that measures attitudes towards mental 
health and frequency of mental health disorders in the tech workplace.

no_employees column have date for some of the rows which is data impurities and 
needs to be cleaned.
  
```{r echo=TRUE, include=TRUE}
setwd("C:/Users/14024/Desktop/dsc520-fork-chitro")

## Load the `2014 data' to 
survey_USA <- read.csv("Final project/survey_USA.csv")

## Load the `2016 data' to 
survey_UK <- read.csv("Final project/survey_UK.csv")

## Load the `Others data' to 
survey_others <- read.csv("Final project/survey_others.csv")
```
## Understanding dataset columns:

* Timestamp : 
* Age : Age of the employee.
* Gender : Male or Female employee.
* Country : Employee working in which country.
* state: If you live in the United States, which state or territory do you live in?
* self_employed: Are you self-employed?
* family_history: Do you have a family history of mental illness?
* treatment: Have you sought treatment for a mental health condition?
* work_interfere: If you have a mental health condition, do you feel that it 
  interferes with your work?
* no_employees: How many employees does your company or organization have?
* remote_work: Do you work remotely (outside of an office) at least 50% of 
  the time?
* tech_company: Is your employer primarily a tech company/organization?
* benefits: Does your employer provide mental health benefits?
* care_options: Do you know the options for mental health care your employer
  provides?
* wellness_program: Has your employer ever discussed mental health as part 
  of an employee wellness program?
* seek_help: Does your employer provide resources to learn more about mental 
  health issues and how to seek help?
* anonymity: Is your anonymity protected if you choose to take advantage of 
  mental health or substance abuse * treatment resources?
* leave: How easy is it for you to take medical leave for a mental health 
  condition?
* mentalhealthconsequence: Do you think that discussing a mental health 
  issue with your employer would have * negative consequences?
* physhealthconsequence: Do you think that discussing a physical health 
  issue with your employer would have * negative consequences?
* coworkers: Would you be willing to discuss a mental health issue with 
  your coworkers?
* physhealthinterview: Would you bring up a physical health issue with 
  a potential employer in an interview?
* mentalvsphysical: Do you feel that your employer takes mental health 
  as seriously as physical health?
* obs_consequence: Have you heard of or observed negative consequences 
  for coworkers with mental health       conditions in your workplace?
* comments: Any additional notes or comments

## Required Packages :

```{r echo=TRUE, include=TRUE}
library("ggplot2")# plot
library("plyr")# data cleanning
library("dplyr")# data cleanning
library("magrittr")
library("foreign")
library("coefplot")
library("caTools")
library("rlang")
library("lazyeval")
library("gridExtra")
library("tidyr")
library("readr")
library("purrr")
library("tibble")
library("stringr")
library("forcats")
library("rworldmap")
##library('maps')# world map
##library("corrplot")# corrplot
##library("fiftystater")# us map
```

## Plots :

* histogram
* boxplot
* Points plot 
* bar plot
* line plot
* Points and line plot 

## Questions for future steps :

* Need to join all 3 datasets to analyze/answer questions across differnt
  countries.
  
* What level of data munjing required to develop a model.

* What should be the split percentage of test and training data for the accuracy 
  testing of the model.
  
## Import and clean data :

* Downloaded the data from kaggle and kept in local machine. Split the 2014
  into 3 different dataset based on the country value. 
* Use read.csv function  to read and import the data into differnt variables.
* Check whcih all columns have 'NA' and count of such rows in each dataset.
* As this data entered in the survey by employee and some of the columns data 
  are not consistent which will be useful for analysis, make themconsistent.
* Remove missing values Sometimes, people tend to not put a value while filling their survey. 
  Since there is an abundance of data instances, it is not too much of a worry 
  to drop them from our model.
* Detect anomalies (outliers) in numeric attributes, in this dataset age is only
  numeric variable.Some of the columns data is invalid which need to filter 
  out before making the dataset final.
  
```{r echo=TRUE, include=TRUE}
setwd("C:/Users/14024/Desktop/dsc520-fork-chitro")

## Load the `USA data' to 
survey_USA <- read.csv("Final project/survey_USA.csv")
#head(survey_2014)
## Load the `UK data' to 
survey_UK <- read.csv("Final project/survey_UK.csv")
#head(survey_UK)
## Load the `Others data' to 
survey_Others <- read.csv("Final project/survey_Others.csv")
#head(survey_Others)

## Display internal structure of survey_USA and Columns contains 'NA' value
str(survey_USA)
sapply(survey_USA, function(x) sum(is.na(x)))

## Display internal structure of survey_UK and Columns contains 'NA' value
str(survey_UK)
sapply(survey_UK, function(x) sum(is.na(x)))

## Display internal structure of survey_Others and Columns contains 'NA' value
str(survey_Others)
sapply(survey_Others, function(x) sum(is.na(x)))

## checking the unique values in the Gender column in each dataset
unique(survey_USA$Gender)
unique(survey_UK$Gender)
unique(survey_Others$Gender)

## converting the column values to lowercase
survey_USA$Gender <- tolower(survey_USA$Gender)
survey_UK$Gender <- tolower(survey_UK$Gender)
survey_Others$Gender <- tolower(survey_Others$Gender)

## defining lists on the basis of unique values present in the dataset
male <- c('male-ish', 'cis male', 'male (cis)', 'make', 'mail', 
          'ostensibly male, unsure what that really means', 'm', 'maile', 
          'male ', 'msle', 'mal', 'man', 'malr', 'cis man', 'male')
female <- c('female', 'female ', 'female (cis)', 'woman', 'cis female', 
            'cis-female/femme', 'femail', 'f', 'femake')
others <- c('p', 'nah', 'all', 'a little about you', 'genderqueer', 
            'non-binary', 'trans woman', 'androgyne', 'neuter', 'trans-female',
            'agender', 'female (trans)', 'something kinda male?', 'enby', 'guy (-ish) ^_^',
            'queer', 'queer/she/they', 'male leaning androgynous', 'fluid', 'genderqueer',
            'non-binary', 'trans woman', 'androgyne', 'neuter', 'trans-female',
            'agender', 'female (trans)', 'something kinda male?', 'enby', 
            'guy (-ish) ^_^', 'queer', 'queer/she/they', 'male leaning androgynous', 'fluid')


## replace the values in the data set so that the categories are consistent
survey_USA <- survey_USA %>% 
  mutate(Gender = replace(Gender, which(Gender %in% male), 'Male'),
                          Gender = replace(Gender, which(Gender %in% female), 'Female'),
                          Gender = replace(Gender, which(Gender %in% others), 'Others'))
survey_UK <- survey_UK %>% 
  mutate(Gender = replace(Gender, which(Gender %in% male), 'Male'),
                          Gender = replace(Gender, which(Gender %in% female), 'Female'),
                          Gender = replace(Gender, which(Gender %in% others), 'Others'))
survey_Others <- survey_Others %>% 
  mutate(Gender = replace(Gender, which(Gender %in% male), 'Male'),
                         Gender = replace(Gender, which(Gender %in% female), 'Female'),
                         Gender = replace(Gender, which(Gender %in% others), 'Others'))
## checking if all values are covered and category is consistent                                                
unique(survey_USA$Gender)
unique(survey_UK$Gender)
unique(survey_Others$Gender)

## checking the summary of the Age column
summary(survey_USA$Age)
summary(survey_UK$Age)
summary(survey_Others$Age)

## age have some negative and more than 100 values which needs to be cleaned as age -ve is
## invalid and more than can be considered as invalid
## keeping rows with age between 18 and 100

survey_USA_df <- survey_USA %>% filter(Age >= 18 & Age < 100)
survey_UK_df <- survey_UK %>% filter(Age >= 18 & Age < 100)
survey_Others_df <- survey_Others %>% filter(Age >= 18 & Age < 100)

survey_USA_final_df <- survey_USA_df %>% select(-c(Timestamp,comments))
survey_UK_final_df <- survey_UK_df %>% select(-c(Timestamp,comments))
survey_Others_final_df <- survey_Others_df %>% select(-c(Timestamp,comments))
```

## What does the final data set look like :
```{r echo=TRUE, include=TRUE}
survey_data <- rbind(survey_USA_final_df, survey_UK_final_df, survey_Others_final_df)
survey_data %>% dim
summary(survey_data)

summarise(survey_data, min_age = min(Age))
summarise(survey_data, max_age = max(Age)) 
summarise(survey_data, med = median(Age)) 

state_freq <- as.data.frame(table(survey_data$state))
colnames(state_freq) <- c("state","freq")
as_tibble(state_freq)
```
## What information is not self-evident :

* This data is entered by employee during survey, so some of the columns such 
  as family_history, mentalhealthconsequence and physhealthconsequence might 
  not be self evident as some of the employees might think they might have 
  some -ve consequnces for providing this info in survey.
  
## Questions for future steps :

* Binary Vairables are treatment, family_history and Tech_company in the dataset
  and No_emloyes is categorical variable. Need to use this for linear and logistic 
  model design.
* Based on which column we should group the data from each dataset which will 
  provide relevant answer to the research questions.
* Identify the variables from dataset which have linear relationship between
  them.
* Identify the variables which will have binomial output and can be used fro 
  logistic regression.
  
## Different ways to look at data and slicing and dicing of data with visualization :
* Visualize based of no_employees in company
* Visualize based on gender
* Visualize based on age
  
```{r echo=TRUE, include=TRUE}
## view the specific number of company in each size range 
survey_data$no_employees %>% table

## ordered no_employees level
  level_no_employees <- factor(survey_data$no_employees, levels = 
                c("1-5","6-25","26-100","100-500","500-1000","More than 1000"))
## view the company distribution based on company size
survey_data %>% ggplot(aes(x=level_no_employees))+
  geom_bar(fill = "green") + ggtitle("ordered company size")

gender_diversity <- survey_data %>%
  group_by(Gender) %>%
  dplyr::summarize(count = n())
gender_diversity


## Visualize the number of subjects in each gender type  
G1 <- survey_data %>% ggplot(aes(x=Gender)) +
  geom_bar() + ggtitle("Categories of Gender")

G2 <- ggplot(gender_diversity, aes(x = Gender, y = count, fill = Gender)) +  
  geom_bar(stat = "identity", alpha = 0.5) +
  xlab("Gender Diversity") + 
  ylab("Number of People") + 
  ggtitle("Comparing Gender Diversity in the 2014 Mental Health in Tech Survey")

grid.arrange(G1, G2, nrow = 1)

#Age distribution
g1 <- survey_data %>% ggplot(aes(x=Age)) +
  geom_histogram() + ggtitle("Distribution of Age")

# Histogram of Age
g2 <- ggplot(survey_data,aes(x=Age))+geom_histogram(aes(y=..density..), fill="green")+geom_density(col="#3438BD",alpha = 0.5)+labs(x="Age",
                       title="Distribution of Transformed Age")
# Age categorization#
survey_data$Age<-cut(survey_data$Age, breaks = c(0, 16, 34, 60, 75), 
                      labels = c('Fresh', 'Junior', 'Senior', 'Super'))

# Verify Age group
survey_data$Age %>% table

# Group by Age Group and count each group
age_group <- survey_data %>%
  group_by(Age) %>%
  dplyr::summarize(count = n())
age_group

g3 <- ggplot(age_group, aes(x = Age, y = count, fill = Age)) +  
  geom_bar(stat = "identity", alpha = 0.5) +
  xlab("Age Group") + 
  ylab("Number of People") + 
  ggtitle("Comparing Age Group in the 2014 Mental Health in Tech Survey")

grid.arrange(g1, g2, g3, nrow = 1)

```
## summarize your data to answer key questions

```{r echo=TRUE, include=TRUE}

# First, re-value the variables of interest
survey_data$no_employees <- as.factor(revalue(survey_data$no_employees,
                       c("1-5"="1", "6-25"="2", "26-100"="3", "100-500"="4",
                         "500-1000"="5", "More than 1000"="6")))

survey_data$treatment <- as.numeric(revalue(survey_data$treatment,
                                           c("No"="0", "Yes"="1")))
# Get a collection of tables for the responses

questions_only <- survey_data[,5:24]

response_tables <- 
  lapply(questions_only, function(Response) as.data.frame(table(Response)))

# Add names of questions for each for plotting purposes
plot_tables <- 
  lapply(seq_along(response_tables), 
         function(x) mutate(response_tables[[x]], name = names(response_tables)[x]))

all_questions <- bind_rows(plot_tables)
all_questions

# Group by family_history and count each group

fh_count <- survey_data %>%
  group_by(family_history) %>%
  dplyr::summarize(count = n(), proportion = n()/nrow(survey_data))
fh_count

# Group by treatment and Gender and count each group

treat_count <- survey_data %>%
  group_by(treatment, Gender) %>%
  dplyr::summarize(count = n(), proportion = n()/nrow(survey_data))
treat_count

# care_options available based on country

care_count <- survey_data %>% filter(care_options == "Yes") %>%
  group_by(Country) %>%
  dplyr::summarize(count = n(), proportion = n()/nrow(survey_data))
care_count


# Group by work_interfere response and count each group

survey_data$work_interfere <- factor(survey_data$work_interfere, ordered = TRUE, 
                  levels = c('NA', 'Never', 'Rarely', 'Sometimes', 'Often'))

wi_count <- survey_data %>%
  group_by(work_interfere) %>%
  dplyr::summarize(count = n(), proportion = n()/nrow(survey_data)) %>%
  arrange(work_interfere)
wi_count

# Group by Country and count each group
country_count <- survey_data %>%
  group_by(Country) %>%
  dplyr::summarize(count = n())
country_count

mental_health_interview <- survey_data %>%
  group_by(mental_health_interview) %>%
  summarise(n = n()) %>%
  mutate( percentage = signif(100 * ( n /sum(n)),2))
mental_health_interview

coworkers <- survey_data %>%
  group_by(coworkers) %>%
  summarise(n = n()) %>%
  mutate( percentage = signif(100 * ( n /sum(n)),2))
coworkers
```

## Ploting to illustrate the finding of key questions :

```{r echo=TRUE, include=TRUE}  
# Does family_history have any relation with mentalhealth

ggplot(fh_count, aes(x = family_history, y = count, fill = family_history)) +  
  geom_bar(stat = "identity", alpha = 0.5) +
  xlab("Family History in Mental illness") + 
  ylab("No of People") + 
  ggtitle("Comparing Family History in the 2014 Mental Health in Tech Survey")

# Use a bar graph to graph `treatment` (treated mental health condition) 
# over `no_employees` (number of employees) and `tech_company` (company type)
ggplot(survey_data,aes(x=no_employees,y=treatment, fill=factor(tech_company)), 
       color=factor(vs)) +  
  stat_summary(fun.y=mean,position=position_dodge(),geom="bar") + 
  labs(x = "Number of employees", y = "Probability of mental health condition", 
       title = "Probability of mental health illness by workplace type and size") +
  scale_x_discrete(labels=c("1" = "1-5", "2" = "6-25", "3" = "26-100", "4"="100-500",
                              "5"="500-1000", "6"=">1000"))
ggplot(fh_count, aes(x = family_history, y = count, fill = family_history)) +  
  geom_bar(stat = "identity", alpha = 0.5) +
  xlab("Family History in Mental illness") + 
  ylab("No of People") + 
  ggtitle("Comparing Family History in the 2014 Mental Health in Tech Survey")

# Visualize the number of subjects   
ggplot(treat_count, aes(x = treatment, y = count, fill = treatment)) +  
  geom_bar(stat = "identity", alpha = 0.5) +
  xlab("Treatment Received in Mental Health Condition") + 
  ylab("No of People") + 
  ggtitle("Comparing the Treatment Received in the 2014 Mental Health in 
          Tech Survey")

# Visualize the number of subjects   
ggplot(wi_count, aes(x = work_interfere, y = count, fill = work_interfere)) +  
  geom_bar(stat = "identity", alpha = 0.5) +
  xlab("The Level of Work Interfere by Mental Health Condition") + 
  ylab("No of People") + 
  ggtitle("Comparing the Level of Work Interfere in the 2014 Mental Health in 
          Tech Survey")

ggplot(data = mental_health_interview) +
  geom_bar(mapping  =  aes(x = "", y = percentage, fill = mental_health_interview),
           stat = "identity", width = 1) +
  geom_text ( aes(x = c(1,1,1.3), y = c(94,44,1),label = paste(percentage,"%")),
              size = 4) +
  ggtitle("Mental Health Interview") +
  labs(x = "Percentage",y = "") +
  guides(fill = guide_legend(title = NULL)) +
  coord_polar("y")

ggplot(data = coworkers) +
  geom_bar(mapping  =  aes(x = "", y = percentage, fill = coworkers), 
           stat = "identity", width = 1) +
  geom_text ( aes(x = c(1,1,1), y = c(89.5,50,8.5),label = paste(percentage,"%")),
              size = 4) +
  ggtitle("Coworkers") +
  labs(x = "Percentage",y = "") +
  guides(fill = guide_legend(title = NULL)) +
  coord_polar("y")

## Few key questions are answered based on the plotting done above and for rest
## of the questions we need to figure out the correlation between those variables 
## and have to design linear model or logistic regression model. 

## Next step we will plan to build a model (Generalized Liner model) and split 
## the input data into train and test data and test the model. Will calculate 
## the accuracy of the model based on the input data.
```
## Final Step :

## The problem statement addressed :    

With the rapid development of tech companies, tech industry 
must exist a little bit or massive pressure on tech employees. At the 
same time, increasingly tech companies start to focus on employees’ 
mental health issues as they want to make sure that the productivity 
of the resource and company doenst get impacted due to 
mental health of the employee. As a part of this project we have considered 
3 data sets of 3 differt countries (US, UK and others) and each of them consits 
27 variables relating to interview questions of mental health care. The main aim 
of this analysis is to focus on the mental halth of employee in tech company, but as 
the dataset contains data about other companies a comparative study can be done 
between tech company and other company.

The goal is to answer the question: In the tech industry, Who needed to seek 
mental health care? Based on attributes: such as treatment, gender, age group, 
family background with mental illness, I want to know which of 
the employees working in the tech industry has the highest frequency 
distribution for seeking mental health care.

## How the problem statement addressed :

To address the problem statement we have cleaned up the data and filtered out
outliers from the dataset and combined all 3 data sets into 1 to perform the 
analysis accross the globe data. Identified the target variable as “treatment” and 
Treatment, Age, Gender, Family_History, no_employees and tech company are the key
variables for this analysis.

treatment, family_history, no_employees, and tech_company variables, 
the pattern and type are so clean that they do not need to inspect and transform.
As the answer of this variables either "Yes" or "No", those ar binary variables.

```{r echo=TRUE, include=TRUE}
# Select variables
df <- survey_data %>% select(treatment, Age, Gender, family_history, no_employees,
                             tech_company) 

# Whether employees have sought mental health care?
df$treatment %>% table

# Whether employees seeking mental health consultation is affected by family history.
df$family_history %>% table

# Initial plan was to filter out the non-tech company data as we were planing to work 
# only on techcompany data, but for now we are just identifying data split between tech 
# and non tech company.

df$tech_company %>% table

# Identified the categorical variables from the dataset to answer some questions.

# The number of employees in a tech company can help me to answer some questions: 
#  Does the size of the company affect the mental health of employees? If the 
#  number of employees is large, does it mean that the competition is stronger
#  than small companies and employees are more likely to have mental health problems?

# view the specific number of company in each size range 
df$no_employees %>% table

# Interpretation of categorical variable

# This variable is categorical; there are six groups.
# Each group represents the number range of employees in a company or organization.
# Based on the summary, there is no missing value under this variable.

# Gender is another categorical variable. We categorised them into 3 categories 
# in the earlier section.

df_gender <- survey_data %>%
mutate(Gender = replace(Gender, which(Gender %in% male), 'Male'),
Gender = replace(Gender, which(Gender %in% female), 'Female'),
Gender = replace(Gender, which(Gender %in% others), 'Others'))

# Verify the result
df_gender$Gender %>% table

# Create the relative frequency table of gender
table(df_gender$Gender)/length(df_gender$Gender)

# Identify the continuous variable from the dataset.

# Age can be used to answer many of the questions from this data and in the eralier
# phse we have filtered out the outliers from the data.
```
## Logistic Regression Modeling

As the topic focuses on the tech field, we have already filtered Tech companies 
at the last assignment. Will use logistic regression to predict the treatment of 
employees who are working in the tech field, so the dataset we used only includes 
tech companies with relevant information. Based on the tech field data summary, 
We can know that 49.9% of employees have sought mental health treatment, 
which means almost half of them meting mental issues before.

```{r echo=TRUE, include=TRUE}  
# Filter a dataset only focusing on tech_company
Tech <- survey_data %>% select(treatment, Age, Gender, family_history, 
            no_employees, tech_company) %>% filter(tech_company == "Yes")
summary(Tech)

# For my logistic model, I pay attention to four predictors I have selected, 
# which are primary elements relating to the mental health treatment.

# Fit logistic model to full data
lm <- glm( treatment ~ Age + Gender + no_employees + family_history, data = 
             Tech, family = "binomial" )
summary(lm)

# observations from summary output :

# According to the summary of the model, we can see that only two critical 
# elements which can affect the treatment that I expected. Those are gender 
# and famly_hisotry since those p-values are smaller than 0.05.

# The model has converted the predictor “Gender” into “Gendermale” and “Genderqueer”:
# this means that the baseline is the “female” level picking from the model 
# automatically and then a dummy variable is defined for “male” to evaluate 
# the effect of being male relative to female. Obviously, the number of 
# females who have sought mental health treatment is higher than males.

# The model has converted the predictor “family_history” into “family_historyYes”:
# this means that the baseline is “Without family history” Level that the model 
# picked automatically and then defined a dummy variable for “family_historyYes” 
# to evaluate the effect of family_history.

# Lets look at the raw logistic regression coefficients from the model
# Raw logistic regression coefficients
  lm %>% coefficients %>% round(3)
  
# look at the odds ratios by exponentiating the log-odds
  lm %>% coefficients %>% exp %>% round(3)
  
# Interpretation of the outcome :

# There are two main effects gendermale and family_history,reporting 
# by percent decreasing or increasing in the odds of treatment.
  
# Age, GenderQueer, no_employees : Not significant.
  
# Family_historyYes : Employees with family mental health history are 5.319 
# times more likely to receive mental health treatment than employees without 
# family history. This result shocked me that the effect of family history is 
# such serious.
```
## Evaluating predictive performance :

To evaluate predective performance, need to split the dataset into test and train
dataset.
Train dataset to train the model and test is to test the accuracy of the model.
Split the train and test data into 80:20 ratio.

```{r echo=TRUE, include=TRUE}
# Randomly select train/test split indices with 80% training size
n <- nrow(Tech)
train_ind <- sample(seq_len(n), size = floor(0.8*n))

# Split training/testing sets
Tech_train <- Tech[train_ind, ]
Tech_test <- Tech[-train_ind, ]

# Define function for feature enginerring pipeline
transformations <- function(Tech) {
   # Gender
   # Create the list of three categories
Male <- c("Male ","Cis Man", "Malr", "Male", "male", "M", "m", "Male-ish", 
          "maile", "Mal", "Male (CIS)", "Cis Male", "Make", "Male", "Man", 
          "msle", "Mail", "cis male")
Female <- c("Female ","femail","Female (cis)","female","Female","F","Woman","f",
            "Femake","woman","Female",
            "cis-female/femme", "Cis Female", "Trans-female", "Female (trans)", 
            "Trans woman")

Queer <-c ("ostensibly male, unsure what that really means",
           "p","A little about you","queer","Neuter","something kinda male?","non-binary",
		"Nah","All","Enby","fluid","Genderqueer","Androgyne","Agender"
           ,"Guy (-ish) ^_^","male leaning androgynous", "queer/she/they")

# Categorize genders
Tech$Gender <- sapply(
  as.vector(Tech$Gender),
  function(x) if(x %in% Male) "Male" else x ) 

Tech$Gender <- sapply(
  as.vector(Tech$Gender),
  function(x) if(x %in% Female) "Female" else x ) 

Tech$Gender <- sapply(
  as.vector(Tech$Gender),
  function(x) if(x %in% Queer) "Queer" else x ) 

# Age
# Replacing negative values and outliers with median
Tech$Age <- as.numeric(Tech$Age)
Tech$Age[which(Tech$Age<0)]<- median(Tech$Age)
Tech$Age[which(Tech$Age>100)]<- median(Tech$Age)

# Summary Age
summary(Tech$Age)

# Age categorization#
Tech$Age1 <- cut(Tech$Age, breaks = c(0, 16, 34, 60, 75), labels = 
                   c('Fresh', 'Junior', 'Senior', 'Super'))

# Verify Age group
Tech$Age1 %>% table

# Return the transformed dataframe
return(Tech)
}
# Apply feature engineering to each subset
Tech_train <- Tech_train %>% transformations
Tech_test <- Tech_test %>% transformations

# Checking train data
Tech_train %>% head(2)

# checking test data
Tech_test %>% head(2)

#use the training set to re-train the logistical regression model and 
#use the trained models to make predictions on the train and test sets.

# Train model
lm_train <- glm(treatment ~ Age + Gender + family_history + no_employees, 
                data = Tech_train, family = "binomial")

# Predict on training set
Tech_train$predict_probs <- predict(lm_train, Tech_train, type = "response")
Tech_train$predict <- ifelse(Tech_train$predict_probs < 0.5, "No", "Yes")

# Predict on test set
Tech_test$predict_probs <- predict(lm_train, Tech_test, type = "response")
Tech_test$predict <- ifelse(Tech_test$predict_probs < 0.5, "No", "Yes")

# Confusion matrix - training data
cm_train <- table(Tech_train$treatment, Tech_train$predict, dnn = c("real", "predict"))
cm_train

accuracy <- (cm_train[[1,1]] + cm_train[[2,2]] ) / sum(cm_train)
accuracy <- accuracy * 100
print(paste(round(accuracy), "%"))

Precision <- cm_train[[2,2]] / sum(cm_train[[2,1]] + cm_train[[2,2]] )
Precision <- Precision * 100
print(paste(round(Precision), "%"))

Recall <- cm_train[[2,2]] / sum(cm_train[[1,2]] + cm_train[[2,2]] )
Recall <- Recall * 100
print(paste(round(Recall), "%"))

# Confusion matrix ~ testing data
cm_test <- table(Tech_test$treatment, Tech_test$predict, dnn = c("real", "predict"))
cm_test

accuracy <- (cm_test[[1,1]] + cm_test[[2,2]] ) / sum(cm_test)
accuracy <- accuracy * 100
print(paste(round(accuracy), "%"))

Precision <- cm_test[[2,2]] / sum(cm_test[[2,1]] + cm_test[[2,2]] )
Precision <- Precision * 100
print(paste(round(Precision), "%"))

Recall <- cm_test[[2,2]] / sum(cm_test[[1,2]] + cm_test[[2,2]] )
Recall <- Recall * 100
print(paste(round(Recall), "%"))

```

## Analysis Outcome : 

1.  "Comparing Family History in the 2014 Mental Health in Tech Survey" plot 
    confirms 61% of employees don't have family history of metal illness among the
    employee participated in the survey.
2.  In tech_company workplace type and size don't have much impact on mental health.
3.  Number of people count is same in case of treatment received in Mental Health
    Condition from employer.
4.  As per the analysis most of the employees thinks sometimes work interfere 
    mental health condition of the employee.
5.  Majority of the employee(80%) don't appear for mental health interview.
6.  62% of employees shares mental health with coworkers.

## Implications :

1. This analysis shows that 61% employees seeking mental health condition don't 
  have any family history of mental illness, so employees shuld not think that
  they might not be having mental health issue as they don't have a family history. 
  If any sort of mental illness they feel, they should go for treatment or discuss 
  with supervisor oe co-worker.
  
2. Intial thought after reviewing the data set was  the larger the number of 
   employees in the tech company, the more fierce the competition, so it means 
   that the number of people seeking psychological stress relief will increase.
   But data show that the number of people seeking mental health treatment has 
   little to do with the size of the company.

3. Having remote work at least 50% of the time versus less than 50% of the 
   time showed no significant difference in mental illness work interference levels.
   
4. Female employees’ mental health in technology industry should be paid more
   attention although the input dataset have majority of male data and % of female data
   analyzed is not significant to draw a fianl conclusion.

5. People feel that their employers somewhat easily sanction leave for 
   mental health issues.The reason maybe that the employer does not want to 
   take any risk of overloading the patient with work.

6. People feel that sharing about their mental or physical health with 
   employers would help them a bit but they are reluctant to share the same 
   with their coworkers.They would prefer to share with only some of the coworkers.

## Limitations:

1.  Majority of males in our data set compared to females and other categories.
2.  Majority data from USA and Canada, so can't draw any conclusion on this topic
    across the globe based on this data.
3.  This Survey data was for 2014, if this data received for multiple years
    then we could have analyzed how the trends looks like year over year.
4.  All employees might not be aware of company provided wellness_program and 
    they might have responded without knowing that.
5.  As this is some sort of personal mental health related data received during
    survey, authenticity of the data might be a question but this might be the 
    best source for this sort of data.

## Concluding Remarks :

1. Mental health of employee is a major factor in any sort of Industry and specifically
   in Tech as in Tech Industry the employee have to do major thinking during any sort 
   of development, testing or support work. Your productivity can't be accurate if your
   mental health is 100%. One of the employees mental health can affect the entire 
   teams and eventually impact a grater audience in the company and that impacts 
   company's productivity. As the analysis shows it's doesn't matter how big 
   the company is, but in current situation to give your 100% output you have
   to have mentally fine.

1. This case study is majorly based on cases in the US and Canada data. so cant 
   draw any concrete conclusion that it's true all across the globe but if this 
   analysis can be done on more data definitely the outcome will be more or less 
   similar which we have identified here.
   
1. It is impossible to draw a specific conclusion from the scale of the company 
   because it depends on the company's work culture, competitiveness between 
   employee, other benefits, type of work etc.

1. Cases show that more than 50% of people surveyed in countries like US,Australia 
   and Canada undergo treatment for mental illness and as it's applicable for 
   1st world country definitely we can conclude it might be similar or worse 
   in 2nd and 3rd world countries.

1. People who are not more prone to work at home are usually bored and 
   filled with anxiety leading to degradation in mental health.  THis analysis 
   was done based on 2014 data and after pandemic more remote work or work from 
   home has been introduced, so definitely the company's should more foucs on
   the employees mental health and bring some plans to keep employees mental health
   better.

1. People who are in the early 30's usually undergo treatment but there
   are extreme cases like 8 years and 72 years people receiving 
   the same treatment. Th   is concludes that people are more competitive in that 
   age and bringing more challenges to reach their goal at that age and eventually
   affecting mental health without.

1. It is interesting to find that people face mental trauma regardless of 
   whether they are self employed or not. Self employed people should have more 
   mental health challenges but that is not the outcome of this analysis.

1. The surveyed people agree that their mental health somewhat affects their 
   productivity at work.

1. To conclude this no one is alone in this World,so share your issues which might
   lead to mental illness with co-workers/family members and take necessary treatment
   on time to make you more productive and achieve your goal.



